---
title: "DataScienceHW5"
author: "Gunes Alkan"
date: "October 14, 2015"
output: 
  html_document: 
    keep_md: yes
---


Now, let's read in data first data set GDP, and manipulate it:

```{r}
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv"
download.file(fileURL, destfile="gdp.csv")
gdp <- read.csv(file="gdp.csv", header=T)
names(gdp)
head(gdp)
#skip first 4 lines
gdp <- read.csv(file="gdp.csv", skip=4) # or the next one works the same
head(gdp)
gdp <- read.csv(file="gdp.csv", skip=5, header=F)
head(gdp)
# say R that when it sees these two .. and Not Available, R will put NA
gdp <- read.csv(file="gdp.csv", skip=5, header=F, na.strings=c("..", "Not Available."))
str(gdp)
#help(read.csv)
gdp <- read.csv(file="gdp.csv", skip=5, header=F, na.strings=c("..", "Not available.", "..Not available."), stringsAsFactors=FALSE)
str(gdp)
#now everthing is in either logical or character not Factor
gdp <- dplyr::select(gdp, V1,V2,V4,V5)
str(gdp)
names(gdp) <- c("CountryCode", "Ranking", "Economy", "GDP")
str(gdp)
#lets change ranking to numeric
gdp$Ranking <- as.numeric(gdp$Ranking)
str(gdp)
summary(gdp)
tail(gdp, 40)
sum(is.na(gdp$Ranking))
dim(gdp)
326-136
summary (gdp[191:326,]) #to make sure it is alll NAs
gdp <- gdp[1:190,]
str(gdp)
gdp2 <- gdp
gdp2$GDP <- as.numeric(gdp2$GDP)
str(gdp2)
#we had all NAs because of the commas between numbers, we need to remove them, try it on your test dataset gdp2
gdp2 <- gdp
gdp2$GDP <- gsub("," , "" ,gdp$GDP)
str(gdp2)
gdp2$GDP <- as.numeric(gdp2$GDP)
str(gdp2)
```

Now, let's manipulate education dataset:

```{r}
#read in data
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv"
download.file(fileURL, destfile="educ.csv")
educ <- read.csv(file="educ.csv", header=T)
names(educ)
str(educ)
#select columns
educ <- dplyr::select(educ, CountryCode, Income.Group)
str(educ)
tail(educ, 40)

```

we see that there are some countries whose income.group is missing such as in 228th observation.So,name them as NAs first:

```{r}
educ2 <- read.csv(file="educ.csv", na.strings=c(""))
educ2 <- dplyr::select(educ2, CountryCode, Income.Group)
# let's prove that it actually named empty ones as NA
educ2[228,]

```

now let's see how many missing ones we have:

```{r}
sum(is.na(educ2))

```
let's make our variables character, instead of factor:

```{r}
educ2$CountryCode <- as.character(educ2$CountryCode)
educ2$Income.Group <- as.character(educ2$Income.Group)
str(educ2)

```

Finally, merging two datasets:

```{r}
MergedData <- merge(x = gdp2, y=educ2, by = "CountryCode", all=TRUE)
names(MergedData)
dim(MergedData)
str(MergedData)

```

###### QUESTION 1 ######
HOW MANY OF THE IDS MATCH?
In order to find the answer, let's check dataset with no missing values

```{r}
### data set with all nonmissing values ###
MergedData2 <- na.omit(MergedData)
str(MergedData2)

```

# Answer 1: 
# Since there is 189 observations, we can clearly say that 189 of IDs matched.

And, in MergedData2 data frame, we have only non-missing values, since we dropped all NAs.


```{r}
#another way of showing this is 
sum(!is.na(unique(MergedData$GDP)))

#also, we can check the number of missings in each variable
sum(!is.na(unique(MergedData$Income.Group)))
sum(!is.na(unique(MergedData$CountryCode)))
sum(!is.na(unique(MergedData$Ranking)))
sum(!is.na(unique(MergedData$Economy)))

#We will continue with using MergedData2 data frame

```

###### QUESTION 2 ######
WHAT IS THE 13TH COUNTRY IN THE RESULTING DATA FRAME?

```{r}
#before sorting it, lets check if everything looks ok:
str(MergedData2)
# SORT DATA FRAME IN ASCENDING ORDER BY GDP RANK
library(dplyr)
MergedData3 <- arrange(MergedData2,GDP)
head(MergedData3)
tail(MergedData3, 10)
#So The United States is the last
head(MergedData3, 14)
```

# Answer 2: 
# 13th country is St. Kitts and Nevis .

###### QUESTION 3 ######
WHAT ARE THE AVERAGE GDP RANKINGS FOR THE "High income: OECD" AND "High income: nonOECD" GROUPS?


```{r}
#MergedData3$Income.Group <- as.factor(MergedData3$Income.Group)
str(MergedData3)
summary(MergedData3)
#finding mean
tapply(MergedData3$Ranking,MergedData3$Income.Group, mean)

```

# Answer 3:
# So Average GDP Ranking For "High income: OECD" : 32.96667  
# and Average GDP Ranking For "High income: nonOECD" : 91.91304 


###### QUESTION 4 ######
CUT THE GDP RANKINGS INTO 5 SEPERATE QUANTILE GROUPS. MAKE A TABLE VERSUS INCOME.GROUP.
HOW MANY COUNTRIES ARE LOWER MIDDLE INCOME BUT AMONG THE 38 NATIONS WITH THE HIGHEST GDP?


```{r}
#Here, we'll use Hmisc package and cut() function to cut it into quatile groups.
library(Hmisc)
MergedData3$RKgr <- cut2(MergedData3$Ranking, g=5)
#change their names
GDPgrLabels <- c("Quantile 1", "Quantile 2", "Quantile 3", "Quantile 4", "Quantile 5")
MergedData3$RKgr <- factor(MergedData3$RKgr, labels=GDPgrLabels)
head(MergedData3)
str(MergedData3)
summary(MergedData3)
table(  MergedData3$RKgr,  MergedData3$Income.Group)

```

# Answer 4: 
# There are 5 Lower middle income countries among the 38 nations with the highest GDP.

















