---
title: "Homework 6"
author: "Gunes Alkan"
date: "November 18, 2015"
output: 
  html_document: 
    keep_md: yes
---


```{r}
#First load required packages
library(class)
library(gmodels)
```

Now, let's read in the data, and see characteristics of variables:


```{r}
esteem <- read.csv("E:/ex1223.csv", header=TRUE)
str(esteem)
esteem$Esteem1 <- as.factor(esteem$Esteem1)
head(esteem)

```

Let's construct a new binary variable from this variable, which takes the value 1 for strong agreement and 0 for agreement, disagreement, or strong disagreement. (combine 4 levels into 2), also transform Income2005 variable:

```{r}
## create a new var
esteem$estnew <- ifelse((esteem$Esteem1 == 1), 1, 0)
esteem$estnew <- as.factor(esteem$estnew)
## log annual income
esteem$Income2005 <- log(esteem$Income2005)

```

We'll now explore the dependence of the probability that a person has positive self-esteem on log annual income (Income2005), intelligence (AFQT), years of education (educ), and gender.

Let's fit logistic model first:

```{r}
## fit logistic model
attach(esteem)
myfit <- glm(estnew ~ Income2005 + AFQT + Educ + Gender, data = esteem, family = "binomial")
summary(myfit)
confint(myfit)  ## adds confidence intervals

```

Looking at the output of logistic procedure, the data provides evidence that at the alpha=0.05 level of significance gender is not a significant estimator (or classifier) of the self-esteem (p-value=0.0932).
The quantitative effects of best classifiers as follows:

It's estimated that each 1 unit increase in log of Income2005 is associated with 18% increase in the odds of the highest self esteem response (95% CI: (7.38, 25.9))

Each one percentile increase in AFQT score is associated with an estimated 0.76% increase in the odds of highest self esteem response. (95% Confidence Interval: (0.40%, 1.12%)).

Associated with each one additional year of education was estimated 7.6% increase in the odds of highest self esteem response( Confidence Interval: 3.5%, 11.7%).

(Although gender is not a good estimator, let's make its interpretation too. After accounting for all other variables, the odds of highest self esteem response were estimated to be 14.6% lower for males than for females.)

(In summary, not all variables are needed to obtain the best classification for this dataset.)

Now let's continue with dividing dataset into two group: training and testing set for our best classifiers:

```{r}
## divide esteem into two groups:test and train
ind <- sample(2, nrow(esteem), replace=TRUE, prob=c(0.67, 0.33))
esteem.train <- esteem[ind==1, c("Income2005" , "AFQT" , "Educ" )]
head(esteem.train)
esteem.test <- esteem[ind==2,c("Income2005" , "AFQT" , "Educ" ) ]
head(esteem.test)

```

Lastly, let's obtain corresponding table :

```{r}
## create labels for the table
esteem.trainLabels <- esteem[ind==1, "estnew"]
esteem.testLabels <- esteem[ind==2, "estnew"]
esteem_pred <- knn(train = esteem.train, test = esteem.test, cl = esteem.trainLabels , k=2)
## plot the table
CrossTable(x = esteem.testLabels, y = esteem_pred, prop.chisq=FALSE)

```






